RationalGRL: A Framework for Argumentation and Goal Modeling

### 2017-08-24 background.tex
Comments
Added small comments in TeX (%FB) and todo's

Figure~\ref{fig:example-small} illustrates a simplified GRL diagram from the traffic simulator design exercise. An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system or the system itself (\texttt{Traffic Simulator}, Figure~\ref{fig:example-small}). Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied. Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way. Softgoals (e.g. \texttt{Improve Simulation}) are often related to non-functional requirements, whereas goals (such as  \texttt{Simulate Traffic}) are related to functional requirements. Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals. In Figure~\ref{fig:example-small}, we have the two tasks \texttt{Static Simulation} and \texttt{Dynamic Simulation}: if the system can perform both a static and a dynamic simulation, it can achieve goal \texttt{Simulate Traffic}. In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g., \texttt{Car Objects}). \todo{Floris}{Sepideh}{Belief elements are not explained and not in the example. I think it would be good to include them, as beliefs allow one to give some sort of rationale for e.g. goals (in fact, one of the reviewers of one of our previous papers argued that beliefs can also be used to capture reasons).}

@@

\begin{enumerate}
\item Will the action realize the desired goal?
\item Are there alternative ways of realizing the same goal or contributing to the same softgoal?
\item Does performing the action have a negative side effect?
\end{enumerate}

@@

@@

@@

Connecting argumentation and goal modelling by providing a mapping between two types of diagram was, in our opinion, ultimately an unsatisfying solution given the problems and requirements described in Section \ref{sect:introduction}. One problem is that the argument diagram is at least as complex as the GRL diagram, so any stakeholder trying to understand the discussion thus far has to parse two complex diagrams containing goals, alternative solutions, tasks, and so forth. Furthermore, in our previous work we did not provide the specific critical questions for goal models (see Section \ref{sect:gmas}). This meant that any counterargument had to be constructed from scratch, as no guidance was given as to possible ways to criticize a GRL model. So the previous iterations of the RationalGRL framework violated requirement 1: argument diagrams do not closely mirror the actual discussions of stakeholders in which ideas are proposed and challenged. Furthermore, not having specific argument schemes and critical questions for goal modelling makes it hard to develop a guiding methodology for the use of argumentation in goal modelling.

>>>>

Figure~\ref{fig:example-small} illustrates a simplified GRL diagram from the traffic simulator design exercise. An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system or the system itself (\texttt{Traffic Simulator}, Figure~\ref{fig:example-small}). Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied. Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way. Softgoals (e.g. \texttt{Improve Simulation}) are often related to non-functional requirements, whereas goals (such as  \texttt{Simulate Traffic}) are related to functional requirements. Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals. In Figure~\ref{fig:example-small}, we have the two tasks \texttt{Static Simulation} and \texttt{Dynamic Simulation}: if the system can perform both a static and a dynamic simulation, it can achieve goal \texttt{Simulate Traffic}.\todo{F}{S}{Later in the paper this goal is called  \texttt{Simulate} , not \texttt{Simulate Traffic}. \texttt{Simulate Traffic} seems more clear to me.} In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g., \texttt{Car Objects}). \todo{F}{S,M}{Belief elements are not explained and not in the example. I think it would be good to include them (or at least discuss them), as beliefs allow one to give some sort of rationale for e.g. goals (in fact, one of the reviewers of one of our previous papers argued that beliefs can also be used to capture reasons).}

@@

%FB: I deleted a few of the critical questions
\begin{enumerate}
\item Will the action realize the desired goal?
\item Are there alternative ways of realizing the same goal or contributing to the same softgoal?
\item Does performing the action have a negative side effect?
\end{enumerate}

@@

%FB: I extended the example and put the figure back in. Note that this is just to explain argumentation, not to reason about the model in 2.2. Putting explicit text in the example makes it more clear.

@@

%FB: I discuss our previous work on RationalGRL here now (used to be in intro), and I discuss why the approach we take in this paper is different from the approach in our previous work.

@@

Connecting argumentation and goal modelling by providing a mapping between two types of diagram was, in our opinion, ultimately an unsatisfying solution given the problems and requirements described in Section \ref{sect:introduction}. One problem is that the argument diagram is at least as complex as the GRL diagram, so any stakeholder trying to understand the discussion thus far has to parse two complex diagrams containing goals, alternative solutions, tasks, and so forth. Furthermore, in our previous work we did not provide the specific critical questions for goal models (see Section \ref{sect:gmas}). This meant that any counterargument had to be constructed from scratch, as no guidance was given as to possible ways to criticize a GRL model. So the previous iterations of the RationalGRL framework violated requirement 1: argument diagrams do not closely mirror the actual discussions of stakeholders in which ideas are proposed and challenged. Furthermore, not having specific argument schemes and critical questions for goal modelling makes it hard to develop a guiding methodology for the use of argumentation in goal modelling (requirement 4).

### 2017-08-24 formalframework.tex
Comments
Added small comments in TeX (%FB) and todo's

@@

\begin{definition}[Argument] \label{def:argument}
An \emph{argument} $A\subseteq Att$ is a set of atoms from $Att$.
\end{definition}

@@

>>>>

\todo{F}{F,M}{Strictly speaking you don't define a set of atoms here, but a set of predicates... A neat solution would be to do it as follows: Def 1 is the \emph{GRL language} L, and we add a new Def for \emph{GRL models} M, which is then a set of atoms based on L. The definition of Argument is then the same as Model, but this is correct (cf. the metamodel: a model is an argument).}

@@

\begin{definition}[Argument] \label{def:argument}
An \emph{argument} $A\subseteq Args$ is a set of atoms from $At$.
\end{definition}

@@

\todo{F}{F,M}{It would be good to have a formal definition of how to construct models given an argumentation framework. If we define a model M at the start of this section, this should be fairly easy.}

### 2017-08-24 gmas.tex
Comments
Added small comments in TeX (%FB) and todo's

\section{Argument Schemes for Goal Modeling}
\label{sect:gmas}

@@

\subsection{Details experiment}

The transcripts we used are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.

>>>>

\section{Argument Schemes for Goal Modeling: a Case Study}
%FB: I added "a case study" to emphasize that we perform a case study on the transcripts (see also the introduction)
\label{sect:gmas}

@@

\subsection{Details Case Study}
\todo{F}{F}{I think it would be better to call what Marc did (i.e. annotating the transcripts with arguments and schemes) a case study instead of an experiment. This also avoids the confusion between our case study (annotating transcripts) and the original experiment. The way it is written now makes it seem like we performed the experiments with the students. I can amend the below text early next week.}
The transcripts we used are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.

### 2017-08-24 introduction.tex
Comments
Added small comments in TeX (%FB) and todo's

@@

@@

\begin{enumerate}
\item 
The argumentation techniques should be close to the actual discussions of stakeholders or designers in the early requirements engineering phase.
\item 
The framework must have formal traceability links between elements of the goal model and underlying arguments.
\item 
Using these traceability links, it must be possible to compute the effect of changes in the goal model on the underlying arguments, and vice versa.
\item 
There should be a methodology for the framework to guide the practitioners in its application in real cases.
\item 
The framework must have software tool support.
\end{enumerate}

Our framework, which we call \emph{RationalGRL}, combines the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} with a technique from argumentation theory and discourse modeling called \emph{argument schemes} (or argumentation schemes~\cite{walton-etal2008}). Argument schemes are reusable patterns of reasoning that capture the typical ways in which humans argue and reason, and they are very well suited for modeling discussions about a goal model, as they can guide users in systematically deriving conclusions and making assumptions explicit~\cite{murukannaiah2015}. One argument scheme that is important when reasoning about goals is the argument scheme for practical reasoning~\cite{walton1990,atkinson2007}, which has been used for capturing multi-agent planning \cite{medellin2013planning}, dialogues about safety critical actions \cite{tolchinsky2012deliberation} and software design discussions~\cite{BlackEtal2013}.

@@

>>>>

%FB: I deleted the subsection headings 1.1, 1.2 because the paper started with heading 1.1 immediately following heading 1. (which is frowned upon). The intro is not overly long now so I don't think we need subsections

@@

%FB: I changed the text above to that below. The above 4 reasons overlap in some cases, I tried to make this more concise.

@@

\begin{enumerate}
\item 
The argumentation techniques should be close to the actual discussions of stakeholders or designers in the early requirements engineering phase.
\item 
The framework must have formal traceability links between elements of the goal model and underlying arguments.
\item 
Using these traceability links, it must be possible to compute the effect of changes in the goal model on the underlying arguments, and vice versa.\todo{F}{M,S}{Given the current logical framework, is it possible to compute effect of changes in GM on ARG? It seems to me that we can only do the other way around with the algorithms.}
\item 
There should be a methodology for the framework to guide the practitioners in its application in real cases.
\item 
The framework must have software tool support.
\end{enumerate}

%In this context, the main research question is:

%\begin{quote}
%\textbf{RQ.} What are the constructs, mechanisms and rules for developing a framework that formally captures the discussions between stakeholders such that it can generate goal models?
%\end{quote}

%FB: I deleted the research question. I think that the requirements already nicely capture exactly what our goal is in the paper. We can rephrase this goal as an RQ, but what does this buy us? Also, the RQ introduces new terminology (construct, mechanisms, rules), to which we have to explicitly refer later in the paper (e.g. we found mechanisms A, B, rules R1, R2, etc.).

To answer the research question above and solve some of the concerns mentioned in Subsection~\ref{sect:problem_statement}, we propose a framework called RationalGRL. The RationalGRL framework combines the Goal-oriented Requirements Language (GRL) with a technique from argumentation theory and discourse modeling called \emph{argument schemes}~\cite{walton-etal2004}. \todo{Floris}{Floris}{mention Singh et al. argument schemes in RE here.}

%FB: Some of the text below explaining our use of argument schemes was taken from 2.4. The discussion of previous work on RationalGRL was moved to 2.4.
Our framework, which we call \emph{RationalGRL}, combines the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} with a technique from argumentation theory and discourse modeling called \emph{argument schemes} (or argumentation schemes~\cite{walton-etal2008}). Argument schemes are reusable patterns of reasoning that capture the typical ways in which humans argue and reason, and they are very well suited for modeling discussions about a goal model, as they can guide users in systematically deriving conclusions and making assumptions explicit~\cite{murukannaiah2015}. One argument scheme that is important when reasoning about goals is the argument scheme for practical reasoning~\cite{walton1990,atkinson2007}, which has been used for capturing multi-agent planning \cite{medellin2013planning}, dialogues about safety critical actions \cite{tolchinsky2012deliberation} and software design discussions~\cite{BlackEtal2013}.

%FB: the below text is largely new/reworked stuff. I link what we do in the paper to each of the 5 requirements

@@

\todo{F}{all}{check text below when we finish the paper}

### 2017-08-24 overview.tex
fixed many things
Comments
Added small comments in TeX (%FB) and todo's

\subsection{RationalGRL Methodology} 
%\todo{Sepideh}{Marc}{Review this section and make sure it is correct}

@@

\begin{figure}[ht]
\centering
\includegraphics[scale=0.4]{img/framework}
\caption{The RationalGRL Framework}
\label{fig:rationalgrl-framework}
\end{figure}

@@

\begin{figure*}[ht]
\centering
\includegraphics[scale=0.4]{img/methodology}
\caption{The RationalGRL Methodology}
\label{fig:rationalgrl-methodology}
\end{figure*}

@@

Figure~\ref{fig:metamodel} depicts the RationalGRL metamodel linking the main elements of our argumentation extension to the main GRL elements. We describe the metamodel bottom-up, starting with the GRL package.

>>>>

\subsection{RationalGRL Methodology}

@@

\begin{figure}[ht]
\centering
\includegraphics[scale=0.4]{img/framework}
\caption{The RationalGRL Framework\todo{F}{M,S}{Can you make this a vector image?}}
\label{fig:rationalgrl-framework}
\end{figure}

@@

\begin{figure*}[ht]
\centering
\includegraphics[scale=0.4]{img/methodology}
\caption{The RationalGRL Methodology\todo{F}{M,S}{Can you make this a vector image?}}
\label{fig:rationalgrl-methodology}
\end{figure*}

@@

Figure~\ref{fig:metamodel} depicts the RationalGRL metamodel linking the main elements of our argumentation extension to the main GRL elements. We describe the metamodel bottom-up, starting with the GRL package.\todo{F}{all}{Fig. 6 is now far removed from where it is mentioned in the text, we should fix this when paper is nearly finished}
